---
title: "Project 2"
author: "Blinda Tian"
date: "11/1/2020"
output: word_document
editor_options: 
  chunk_output_type: inline
---

# Open File
datainput 11 days

```{r}
library(jsonlite)

dat_1 <- read_json('~/Downloads/gps/20200818114606.geojson')
dat_1 <- dat_1$features
dat_2 <- read_json('~/Downloads/gps/20200819132607.geojson')
dat_2 <- dat_2$features
dat_3 <- read_json('~/Downloads/gps/20200820151044.geojson')
dat_3 <- dat_3$features
dat_4 <- read_json('~/Downloads/gps/20200821111447.geojson')
dat_4 <- dat_4$features
dat_5 <- read_json('~/Downloads/gps/20200824130857.geojson')
dat_5 <- dat_5$features
dat_6 <- read_json('~/Downloads/gps/20200825121346.geojson')
dat_6 <- dat_6$features
dat_7 <- read_json('~/Downloads/gps/20200826131614.geojson') 
dat_7 <- dat_7$features
dat_8 <- read_json('~/Downloads/gps/20200827113234.geojson')
dat_8 <- dat_8$features
dat_9 <- read_json('~/Downloads/gps/20200828122627.geojson')
dat_9 <- dat_9$features
dat_10 <- read_json('~/Downloads/gps/20200828130816.geojson')
dat_10 <- dat_10$features
dat_11 <- read_json('~/Downloads/gps/20200831115147.geojson')
dat_11 <- dat_11$features

```



# Open and Conveetion  data
1. variables in file include longitude & latitude , time(UTM), distance, time_long, accuracy, altitude, bearing, speed
2. function to grab time, longitude, latitude
3. function to remove dupliacates, then reorder time 
4. function to covert UTM and timestamp
5. function to distance matrix between each stops
6. function to time difference (between two records) timetravel from start time, all in secs and speed
7  add days number (infact  data 9 and data 10 is on the Same day)


```{r setup, include=FALSE}


# variables in file
head(dat_1[['features']], 1) # variables: longitude & latitude , time(UTM), distance, time_long, accuracy, altitude, bearing, speed



# function to grab time, longitude, latitude, speed
data_grab <- function(dat_1){
  time <- c()
  latitude <- c()
  longitude <- c()
  for (i in 1:length(dat_1)){
  time[i] <- dat_1[[i]][["properties"]][["time"]]
  latitude[i]  <- dat_1[[i]][["geometry"]][["coordinates"]][[2]]
  longitude[i]  <- dat_1[[i]][["geometry"]][["coordinates"]][[1]]
  }
  dat_1 <- data.frame(time, latitude, longitude)
  return(dat_1) 
}

dat_1 <- data_grab(dat_1)
dat_2 <- data_grab(dat_2)
dat_3 <- data_grab(dat_3)
dat_4 <- data_grab(dat_4)
dat_5 <- data_grab(dat_5)
dat_6 <- data_grab(dat_6)
dat_7 <- data_grab(dat_7)
dat_8 <- data_grab(dat_8)
dat_9 <- data_grab(dat_9)
dat_10 <- data_grab(dat_10)
dat_11 <- data_grab(dat_11)


# function to remove dupliacates and reorder time 
library(tidyverse)

duplicate_out <- function(dat_1){

dat_1 <- dat_1 %>% distinct(dat_1$longitude,dat_1$latitude, .keep_all = TRUE)
dat_1 <- dat_1[,c("time","longitude","latitude")]
dat_1 <- dat_1[order(dat_1$time),] # reordered based on time
return (dat_1)
}

dat_1 <- duplicate_out(dat_1)
dat_2 <- duplicate_out(dat_2)
dat_3 <- duplicate_out(dat_3)
dat_4 <- duplicate_out(dat_4)
dat_5 <- duplicate_out(dat_5)
dat_6 <- duplicate_out(dat_6)
dat_7 <- duplicate_out(dat_7)
dat_8 <- duplicate_out(dat_8)
dat_9 <- duplicate_out(dat_9)
dat_10 <- duplicate_out(dat_10)
dat_11 <- duplicate_out(dat_11)


# function to covert UTM and time
library("sp")
convert_umt <- function(dat_1){
 
  spat_df <- c()
  spat_df <- SpatialPointsDataFrame(coords=dat_1[, c("longitude", "latitude")],
                                  data=dat_1['time'],   # This needs to be a data frame
                                  proj4string=CRS("+proj=longlat +datum=WGS84")) 
  
  
  utm_df <- c()
  utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84") 
  utm_coords <- coordinates(utm_df)
  dat_1["UTM"] <- utm_coords
  return(dat_1)
}

dat_1<- convert_umt(dat_1)
dat_2 <- convert_umt(dat_2)
dat_3 <- convert_umt(dat_3)
dat_4 <- convert_umt(dat_4)
dat_5 <- convert_umt(dat_5)
dat_6 <- convert_umt(dat_6)
dat_7 <- convert_umt(dat_7)
dat_8 <- convert_umt(dat_8)
dat_9 <- convert_umt(dat_9)
dat_10 <- convert_umt(dat_10)
dat_11 <- convert_umt(dat_11)



# function to distance matrix between each stops
library("fields")
library("Imap")


distance <- function(dat_1){
dist_dat1 <- sp::spDists(dat_1)
dim(dist_dat1)
dist_dat1[1:5, 1:5] 

dat_1["Distance"]<- c(NA, dist_dat1[row(dist_dat1) == col(dist_dat1) + 1]) # distance between two spots over time in meters

return(dat_1)
}
dat_1 <- distance(dat_1)
dat_2 <- distance(dat_2)
dat_3 <- distance(dat_3)
dat_4 <- distance(dat_4)
dat_5 <- distance(dat_5)
dat_6 <- distance(dat_6)
dat_7 <- distance(dat_7)
dat_8 <- distance(dat_8)
dat_9 <- distance(dat_9)
dat_10 <- distance(dat_10)
dat_11 <- distance(dat_11)




# function to time difference (between two records) timetravel from start time, all in secs and speed

time_diff <- function(dat_1){
time_1  <- dat_1$time
time_1 <- as.POSIXct(time_1, tz = "UTC", "%Y-%m-%dT%H:%M:%OS")
diff_time_1 <- c()
Timetravel <- c()

for(i in 1:length(time_1)){
 Timetravel[i] <- difftime( time_1[i], time_1[1],units = c("secs"))}

for(i in 1:length(time_1)-1){
  diff_time_1[i+1] <- Timetravel[i+1]-Timetravel[i]
  
}
dat_1["Time_diff"] <- c(diff_time_1)
dat_1["Speed"]<- c( dat_1$Distance/dat_1$Time_diff)
dat_1["Timetravel"] <- Timetravel
return(dat_1)
}

dat_1 <- time_diff(dat_1)
dat_2 <- time_diff(dat_2)
dat_3 <- time_diff(dat_3)
dat_4 <- time_diff(dat_4)
dat_5 <- time_diff(dat_5)
dat_6 <- time_diff(dat_6)
dat_7 <- time_diff(dat_7)
dat_8 <- time_diff(dat_8)
dat_9 <- time_diff(dat_9)
dat_10 <- time_diff(dat_10)
dat_11 <- time_diff(dat_11)









# add days number
dat_1["Day"] <- 1
dat_2["Day"] <- 2
dat_3["Day"] <- 3
dat_4["Day"] <- 4
dat_5["Day"] <- 5
dat_6["Day"] <- 6
dat_7["Day"] <- 7
dat_8["Day"] <- 8
dat_9["Day"] <- 9
dat_10["Day"] <- 9  # data 9 and data 10 Same day
dat_11["Day"] <- 11


```












# Data Summarization
Timestamp_count
Duration
Start
Pause
Slocal
Plocal
Slocal
Plocal
startlong
startlat

```{r}


# data summarization 11days 
dat_sum <- list(dat_1,dat_2,dat_3,dat_4,dat_5,dat_6,dat_7,dat_8,dat_9,dat_10,dat_11)

dat_summarize <- c()
Timestamp_count <- c()
Duration <-c()
for(i in 1:length(dat_sum)){
  Timestamp_count[i] <- length(dat_sum[[i]][["time"]])
  Duration[i] <- sum(dat_sum[[i]]$Time_diff,na.rm=TRUE)
}
data_summarize <- data.frame(Timestamp_count,Duration)
data_summarize


# Time Start and Pause

library("countytimezones")
summary_grab <- function(dat_1){
  dat1_s <- format(as.POSIXct(dat_1$time[1], tz = "UTC","%Y-%m-%dT%H:%M:%OS"),'%A, %B %d, %Y %H:%M:%S') #to UTM
  local_t <- calc_single_datetime(as.POSIXct(dat_1$time[1], tz = "UTC","%Y-%m-%dT%H:%M:%OS"), tz = "MST") # to local time (MST)
  dat1_p <- format(as.POSIXct(tail(dat_1$time,1), tz = "UTC","%Y-%m-%dT%H:%M:%OS"),'%A, %B %d, %Y %H:%M:%S') #to UTM 
  local_p<- calc_single_datetime(as.POSIXct(tail(dat_1$time,1), tz = "UTC","%Y-%m-%dT%H:%M:%OS"), tz = "MST") # to local time (MST)

# start location and stop location
  long_s <- (dat_1$longitude[1])
  long_p <- tail(dat_1$longitude,1)
  lat_s <- (dat_1$latitude[1])
  lat_p <- tail(dat_1$latitude,1)
  # return 
    return(c(dat1_s,local_t,dat1_p,local_p,long_s,lat_s,long_p,lat_p))
}

dat1_s<- summary_grab(dat_1)
dat2_s<- summary_grab(dat_2)
dat3_s<- summary_grab(dat_3)
dat4_s<- summary_grab(dat_4)
dat5_s<- summary_grab(dat_5)
dat6_s<- summary_grab(dat_6)
dat7_s<- summary_grab(dat_7)
dat8_s<- summary_grab(dat_8)
dat9_s<- summary_grab(dat_9)
dat10_s<- summary_grab(dat_10)
dat11_s<- summary_grab(dat_11)



data_summarize$Start <- c(dat1_s[1],dat2_s[1],dat3_s[1],dat4_s[1],dat5_s[1],dat6_s[1],dat7_s[1],dat8_s[1],dat9_s[1],dat10_s[1],dat11_s[1])
data_summarize$Slocal <- c(dat1_s[2],dat2_s[2],dat3_s[2],dat4_s[2],dat5_s[2],dat6_s[2],dat7_s[2],dat8_s[2],dat9_s[2],dat10_s[2],dat11_s[2])
data_summarize$Pause<- c(dat1_s[3],dat2_s[3],dat3_s[3],dat4_s[3],dat5_s[3],dat6_s[3],dat7_s[3],dat8_s[3],dat9_s[3],dat10_s[3],dat11_s[3])
data_summarize$Plocal <- c(dat1_s[4],dat2_s[4],dat3_s[4],dat4_s[4],dat5_s[4],dat6_s[4],dat7_s[4],dat8_s[4],dat9_s[4],dat10_s[4],dat11_s[4])

data_summarize$startlong <-c(dat1_s[5],dat2_s[5],dat3_s[5],dat4_s[5],dat5_s[5],dat6_s[5],dat7_s[5],dat8_s[5],dat9_s[5],dat10_s[5],dat11_s[5])

data_summarize$startlat <-c(dat1_s[6],dat2_s[6],dat3_s[6],dat4_s[6],dat5_s[6],dat6_s[6],dat7_s[6],dat8_s[6],dat9_s[6],dat10_s[6],dat11_s[6])
data_summarize


```










# Description of data with visualization
1. plot locations in the past two weeks (days diff cols )
2. spped over time 




```{r}

dat_sum2 <- rbind(dat_1,dat_2,dat_3,dat_4,dat_5,dat_6,dat_7,dat_8,dat_9,dat_10,dat_11)

# travel distance from the initial coordinates 



# All records in 10days
library(ggplot2)
qplot(dat_sum2$longitude, dat_sum2$latitude, colour =  dat_sum2$Day,
      xlab="Longitude", ylab="Latitude", main="Coordinates in two weeks data")



# Zoom in to look at records, without 8/27 coordinates
library(ggplot2)
qplot(dat_sum2$longitude[dat_sum2$Day!=7],
      dat_sum2$latitude[dat_sum2$Day!=7],
      colour=dat_sum2$Day[dat_sum2$Day!=7],
      xlab="Longitude",ylab="Latitude", main="Without 8/27")




# Speed 
par(2,4)

hist(dat_sum2$Speed)
boxplot(dat_sum2$Speed)
  # take out stationary (<0),  extreme value (>30), and target speed will be 1-7
  #placed within 5 meters and 10 seconds to my position --> target walking & cycling speeed


dat_sum3 <-dat_sum2[!dat_sum2$Speed>=7,] 
dat_sum3 <- dat_sum3[!dat_sum3$Speed<=1,]

library(ggplot2)
p <- dat_sum3 %>%
  ggplot( aes(x=Timetravel, y=speed)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Speed Over Time") 

plot(dat_sum3$Speed, dat_sum3$Timetravel)



#Bomb Place : 




```


# Model
Bomb Place : 
placed within 5 meters and 10 seconds to my position --> target walking & cycling speeed

take out 8/27 
```{r}


# take out day 7
dat_sum3 <-dat_sum3[!dat_sum3$Day==7,] 

dim(dat_sum3) # 2923  points


#
m1 <- lm( latitude ~ longitude + Timetravel , data = dat_sum3)
m1_predict <- predict(m1, newdata = dat_sum3, interval = "confidence")


dat_sum3$y_hat <- m1_predict[,1]
dat_sum3$y_lwr <- m1_predict[,2]
dat_sum3$y_upr <- m1_predict[,3]

plot(dat_sum3$longitude,dat_sum3$y_hat )

ggplot(data=dat_sum3, aes(longitude, latitude)) + 
  geom_point(col="steelblue", size=2) + 
  geom_line(aes(longitude, y_hat), col="red") +
  geom_ribbon(aes(ymin=y_lwr, ymax=y_upr), fill="magenta", alpha=.25) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"))


plot(m1$residuals)
acf(m1$residuals)
# The Deviance Residuals: shows us the variation in how far away our observations are from the predicted values. Observations with a deviance residual in excess of two may indicate lack of fit.

summary.lm(m1)



par(mfrow = c(2, 2))
plot(m1)


# check whether there is any residual spatial autocorrelation?



```








# Model validation
Cross Validation
validate the model with some independent data.
Typically, we do this by removing a 20%  of the data to act as a validation dataset, fitting the model on the remaining 80% and then predicting to the validation set. 

```{r}
# take 20% to act as validation set
dim(dat_sum3)
set.seed(1)
validation_rows <- sample(1:nrow(dat_sum3), 2923 *0.2)
dat_sum3_train <- dat_sum3[-validation_rows,] # 80%
dat_sum3_valid <- dat_sum3[validation_rows,] #20%

# Fit model using 80%


m1_validation <- lm( latitude ~ longitude + Timetravel , data = dat_sum3_train)

predictions_validation <- -1.324103e+02-1.572759e+00*dat_sum3_valid$longitude+5.623e-08  *dat_sum3_valid$Timetravel


# Calculate mse
sqrt( mean( (dat_sum3_valid$latitude-predictions_validation)^2 , na.rm = TRUE ) )


#mse(dat_sum3_valid$latitude,predictions_validation,na.rm = TRUE)
ggplot() + geom_point(aes(dat_sum3_valid$longitude,predictions_validation),col="red") + geom_point(aes( dat_sum3_valid$longitude,dat_sum3_valid$latitude),col="Blue") 




```


